---
title: "World Fairs and CO2 Emissions"
filters:
  - shinylive
author: "Shirley Toribio"
format: html
editor: visual
---

```{r, include=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(janitor)
```

## Analyzing data on world fairs from 1855 to 2022

This interactive graph looks at world fairs. According to Wikipedia,

> *a world's fair, also known as a universal exhibition or an expo, is a large global exhibition designed to showcase the achievements of nations. These exhibitions vary in character and are held in different parts of the world at a specific site for a period of time, typically between three and six months.*

Due to various factors, however, it is possible that world fairs might not be as representative of nations worldwide as their description and name would suggest. Ideally, as many nations as possible should be included in a world fair in order for it to be truly representative. As such, I track the number of countries that have attended world fairs from 1851 to 2021 to see how world fairs have evolved in terms of country representation.

```{r, include=FALSE, warning=FALSE, message=FALSE}
setwd(getwd())
fair_data <- read_csv("worlds_fairs.csv") %>%
  mutate(fair_length = case_when(
    start_year == end_year ~ end_month - start_month,
    start_year != end_year ~ (12-start_month) + end_month
  ))
```

```{shinylive-r}
#| fig.alt: "A scatter plot titled number of countries attending world fair based on year. Each point represents a specific world fair. The x-axis is year that fair started and the y-axis is number of countries attending. There is a line of best fit that trends upwards as year increases."
#| standalone: true
#| viewerHeight: 600
library(shiny)
library(bslib)
library(tidyverse)

fair_data <- read_csv("worlds_fairs.csv") %>%
  mutate(fair_length = case_when(
    start_year == end_year ~ end_month - start_month,
    start_year != end_year ~ (12-start_month) + end_month
  ))

ui <- page_sidebar(theme = theme,
                   sidebar = sidebar(
                     sliderInput("rng", "X-Value Range", value = c(1900, 1910), min = min(fair_data$start_year), max = max(fair_data$start_year), sep="")
                   ),
                   mainPanel(plotOutput("plot", width=600))
)

server <- function(input, output, session){
  output$plot <- renderPlot({
    ggplot(fair_data, aes(x = start_year, y = attending_countries)) +
      geom_point(color = "purple") +
      geom_smooth(se=FALSE,color="blue") +
        labs(
          x = "Year that fair started",
          y = "Number of countries attending",
          title = "Number of countries attending world fair based on year"
        ) +
      theme_minimal() +
      xlim(input$rng[1],input$rng[2])
  })
}

shinyApp(ui, server)

## file: worlds_fairs.csv
{{< include worlds_fairs.csv >}}
```

Based on the upward trend of the line of best fit, world fairs seem to have become more inclusive of countries as time has gone by. This trend towards increased inclusivity of countries seems to be especially present from 1975 to 2021. Modern world fair thus seem more representative of countries worldwide than world fairs of the past.

Data source [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2024/2024-08-13).

## Total carbon dioxide emissions from major natural resource companies

This interactive graph looks at emissions of carbon dioxide from major natural resource companies. The dataset used for this analysis originates from a [Carbon Majors](https://carbonmajors.org) [database](https://carbonmajors.org/Downloads).

```{shinylive-r}
#| fig.alt: "A line plot titled total emissions of co2 based on source. The x-axis is year and the y-axis is total emissions of co2 in million tonnes. There are nine lines on the plot and each line represents a different commodity. The nine commodoties represented are anthacite coal, bitumunous coal, cement, lignite coal, metallurgical coal, natural gas, oil, sub-bituminuous coal, and thermal coal. After the year 1950, the lines for oil, bituminuous coal, and natural gas are all higher than the other lines. After 1950, the line for oil is the highest, the line for bituminous coal is the second highest, and the line for natural gas is the third highest."
#| standalone: true
#| viewerHeight: 600
library(shiny)
library(bslib)
library(tidyverse)

emissions_data <- read_csv("emissions.csv") %>% 
  janitor::clean_names() %>%
  group_by(commodity, year) %>%
  summarize(total_emissions = sum(total_emissions_mt_co2e))

commodities <- unique(emissions_data$commodity)

ui <- page_sidebar(theme = theme,
                   sidebar = sidebar(
                     checkboxGroupInput("commodit", "Choose which commodities should be represented in the plot", commodities)
                   ),
                   mainPanel(plotOutput("plot", width=600))
)

server <- function(input, output, session){
  output$plot <- renderPlot({
    data <- emissions_data %>%
      filter(commodity %in% input$commodit)
    ggplot(data, aes(x = year, y = total_emissions)) +
      geom_smooth(aes(color = commodity),se=FALSE) +
      theme_minimal() +
      labs(
        x = "Year",
        y = "Total emissions of co2 in million tonnes ",
        title = "Total emissions of co2 based on source"
      )
  })
}

shinyApp(ui, server)

## file: emissions.csv
{{< include emissions.csv >}}
```

Based on this interactive graph, the production of oil, natural gas, and coal seems to be mostly responsible for the amount of carbon dioxide emitted by major natural resource companies. Major natural resource companies should thus look into reducing their production of these commodities and shifting towards production of other commodoties in order to reduce their carbon dioxide emissions.

Data source [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2024/2024-05-21).
