---
title: "SQL"
author: "Shirley Toribio"
format: html
editor: visual
---

I plan to query data from the Wideband Acoustic Immittance Database, which is a repository of auditory measurements from different people. From this data, two graphs will be generated. One graph will replicate Figure 1 from Voss(2020) and the other will compare mean absorbance measurements between people of different sexes over different frequencies.

```{r, warning=FALSE, message=FALSE}
library(RMariaDB)
library(dbplyr)
library(dplyr)
library(tidyverse)
```

```{r}
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

 #collect(Measurements)
```

You can add options to executable code like this

```{sql}
#| connection: con_wai
#| output.var: "table1"
SELECT Measurements.Identifier,
       Measurements.Frequency,
       AVG(Absorbance) AS mean_absorbance,
       PI_Info.AuthorsShortList AS authors
FROM Measurements
JOIN PI_Info ON Measurements.Identifier = PI_Info.Identifier
GROUP BY Measurements.Identifier, Measurements.Frequency;
```

```{r, warning=FALSE, message=FALSE}
table1 %>%
  filter(authors %in% c("Abur et al.", "Feeney et al.", "Groon et al.", "Lewis and Neely", "Liu et al.", "Rosowski et al.", "Shahnaz et al.", "Shaver and Sun", "Sun et al.", "Voss and Allen", "Voss et al.", "Werner et al.")) %>%
  ggplot(aes(x = Frequency, y = mean_absorbance)) +
  geom_smooth(aes(color = authors), se = FALSE) +
  xlim(0, 8000)
```

The `echo: false` option disables the printing of code (only output is displayed).

```{sql}
#| connection: con_wai
#| output.var: "table2"
SELECT
  Subjects.Sex AS sex,
  Subjects.Race AS race,
  Subjects.Ethnicity AS ethnicity,
  Subjects.Identifier,
  Measurements.Identifier,
  Measurements.Frequency AS freq,
  AVG(Measurements.Absorbance) AS mean_absorbance
FROM Subjects
JOIN Measurements ON Subjects.SubjectNumber = Measurements.SubjectNumber 
WHERE Subjects.Identifier = "Aithal_2013" AND Measurements.Identifier = "Aithal_2013"
GROUP BY ethnicity, race, sex, freq;




```

```{r, warning=FALSE, message=FALSE}
table2 <- table2[,-c(4)]
table2 %>%
  ggplot(aes(x = freq, y = mean_absorbance, color = sex)) +
  geom_smooth(se = FALSE)

```

Using SQL queries, I filtered through and sorted data in a manner that allowed me to compare absorbance measurements across sexes and copy figure 1 of Voss(2020). I did this with 2 SQL queries and joined different tables from the same database to produce both graphs. Following each query, I used ggplot to plot mean absorbance measurements alongside frequency.
