---
title: "Wideband Acoustic Immittance Data Visualization"
filters:
  - shinylive
author: "Shirley Toribio"
format: html
editor: visual
---

## Analysis

I plan to query data from the Wideband Acoustic Immittance Database, which is a repository of auditory measurements from different people. From this data, two graphs will be generated. One graph will be an interactive Shiny version of Figure 1 from Voss([2019](https://journals.lww.com/ear-hearing/citation/2019/11000/resource_review.21.aspx)) and the other will compare mean absorbance measurements between babies of different sexes over different frequencies using data from Aithal et al.([2013](https://www.sciencedirect.com/science/article/pii/S0165587612005332?via%3Dihub)).

This is the abstract of Aithal et al.(2013):

> *Presently, normative wideband reflectance data are available for neonates who have passed a distortion product otoacoustic emission test. However, passing the distortion product otoacoustic emission test alone does not ensure normal middle ear function. The objective of this study was to establish normative wideband reflectance data in healthy neonates with normal middle ear function, as justified by passing a battery of tests.*

Voss(2019) is a summarization of different studies analyzing auditory measurements among people.

```{r, warning=FALSE, message=FALSE}
library(RMariaDB)
library(dbplyr)
library(dplyr)
library(tidyverse)
```

```{r}
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

```

```{sql}
#| connection: con_wai
#| output.var: "table1"
SELECT Measurements.Identifier,
       Measurements.Frequency,
       AVG(Absorbance) AS mean_absorbance,
       CONCAT(PI_Info.AuthorsShortList," (", PI_Info.Year, ")"," N=", COUNT(DISTINCT Measurements.SubjectNumber) ,"; ",Measurements.Instrument) AS studies
FROM Measurements
JOIN PI_Info ON Measurements.Identifier = PI_Info.Identifier
WHERE PI_Info.AuthorsShortList IN ("Abur et al.", "Feeney et al.", "Groon et al.", "Lewis and Neely", "Liu et al.", "Rosowski et al.", "Shahnaz et al.", "Shaver and Sun", "Sun et al.", "Voss and Allen", "Voss et al.", "Werner et al.") AND Measurements.Frequency BETWEEN 0 and 8000
GROUP BY Measurements.Identifier, Measurements.Frequency,  Measurements.Instrument;

```

```{r}
# Code to convert SQL table to csv file that can be used in Shiny app
tf <- cat(NULL,file="tf.csv")
write.csv(table1, file="tf.csv")
```

```{shinylive-r}
#| fig.alt: "A line plot titled mean absorbance from publications in WAI database. Each line represents data from a specific study in the WAI database. There are twelve lines representing data from twelve studies. The x-axis is frequency and the y-axis is mean absorbance. The lines trend upwards from frequencies 0 to 3000 and trend downwards from frequencies 3000 to 6000."
#| standalone: true
#| viewerHeight: 600
library(shiny)
library(bslib)
library(tidyverse)

table1 <- read_csv("tf.csv") 
studies <- unique(table1$studies)

ui <- page_sidebar(theme = theme,
                   sidebar = sidebar(
                     checkboxGroupInput("studies", "Choose which studies should be represented in the plot", studies)
                   ),
                   mainPanel(plotOutput("plot", width=600))
)

server <- function(input, output, session){
  output$plot <- renderPlot({
  data <- table1 %>%
    filter(studies %in% input$studies)
  data %>%
    ggplot(aes(x = Frequency, y = mean_absorbance)) +
    geom_line(aes(color = studies), se = FALSE) + 
    labs(
      title = "Mean absorbance from publications in WAI database"
    )
  })
}

shinyApp(ui, server)

## file: tf.csv
{{< include tf.csv >}}
```


```{sql}
#| connection: con_wai
#| output.var: "table2"
SELECT
  Subjects.Sex AS sex,
  Subjects.Race AS race,
  Subjects.Ethnicity AS ethnicity,
  Subjects.Identifier,
  Measurements.Frequency AS freq,
  AVG(Measurements.Absorbance) AS mean_absorbance
FROM Subjects
JOIN Measurements ON Subjects.SubjectNumber = Measurements.SubjectNumber 
WHERE Subjects.Identifier = "Aithal_2013" AND Measurements.Identifier = "Aithal_2013"
GROUP BY ethnicity, race, sex, freq;
```

```{r}
# Code to convert SQL table to csv file that can be used in Shiny app
df <- cat(NULL,file="df.csv")
write.csv(table2, file="df.csv")
```

```{shinylive-r}
#| standalone: true
#| viewerHeight: 600
library(shiny)
library(bslib)
library(tidyverse)

table2 <- read_csv("df.csv") 
options <- c("nothing","sex")

ui <- page_sidebar(theme = theme,
                   sidebar = sidebar(
                     selectInput("option", "Compare mean absorbances of babies based on", options),
                     sliderInput("rng", "frequency range", value = c(2000, 4000), min = min(table2$freq), max = max(table2$freq), sep="")
                   ),
                   mainPanel(plotOutput("plot", width=600))
)

server <- function(input, output, session){
  output$plot <- renderPlot({
    if (input$option == "nothing"){
      #| fig.alt: "A line plot titled mean absorbance from babies. There is one line representing all babies. The x-axis is frequency and the y-axis is mean absorbance. "
      table2 %>%
        ggplot(aes(x = freq, y = mean_absorbance)) +
        geom_point(color = "purple") +
        geom_line(se = FALSE) +
        labs(
          title = "Mean absorbance from babies"
        )  +
      xlim(input$rng[1],input$rng[2])
    } else if (input$option == "sex"){
      #| fig.alt: "A line plot titled mean absorbance from babies of different sexes. Each line represents data of  babies of a specific sex. There are two lines representing male babies and female babies. The x-axis is frequency and the y-axis is mean absorbance. The line representing male babies tends to be above the line representing female babies."
      table2 %>%
        ggplot(aes(x = freq, y = mean_absorbance, color = sex)) +
        geom_point() +
        geom_line(se = FALSE) +
        labs(
          title = "Mean absorbance from babies of different sexes"
        )  +
      xlim(input$rng[1],input$rng[2])
    }
  })
}

shinyApp(ui, server)

## file: df.csv
{{< include df.csv >}}
```

```{r, warning=FALSE, message=FALSE}
#| fig.alt: "A line plot titled mean absorbance from babies of different sexes. Each line represents data of  babies of a specific sex. There are two lines representing male babies and female babies. The x-axis is frequency and the y-axis is mean absorbance. The line representing male babies tends to be above the line representing female babies."

table2 %>%
  ggplot(aes(x = freq, y = mean_absorbance, color = sex)) +
  geom_line(se = FALSE) +
  labs(
    title = "Mean absorbance from babies of different sexes"
  )
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
dbDisconnect(con_wai)
```

Using SQL queries, I filtered through and sorted data in a manner that allowed me to compare absorbance measurements across babies of different sexes with data from Aithal([2013]())and copy figure 1 of Voss(2019). I did this with 2 SQL queries and joined different tables from the same database to produce both graphs. Following each query, I used ggplot to plot mean absorbance measurements alongside frequency.

### References

Aithal et al. 2013. "Normative wideband reflectance measures in healthy neonates." International Journal of Pediatric Otorhinolaryngology 77 (1). https://doi.org/10.1016/j.ijporl.2012.09.02

Voss, SE. 2019. “Resource Review.” Ear and Hearing 40 (6). https://doi.org/10.1097/AUD.0000000000000790.
